---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
if(!file.exists("activity.csv")){
    unzip("activity.zip")
}
rawData <- read.csv("activity.csv")
ActivityData <- rawData[complete.cases(rawData),]

```



## What is mean total number of steps taken per day?
let's firstly take a look at the chart.
```{r}
library(dplyr)
library(ggplot2)

result <- aggregate(ActivityData$steps, by = list(ActivityData$date), FUN=sum)
colnames(result)<- c("date","steps")
barplot(result$steps, xlab="date", ylab="total steps", space=0, names.arg = result$date, main = "Total number of steps taken each day")

```

let's take a look at the mean and median of the total number of steps taken per day. 
```{r}
summary(result)
```

## What is the average daily activity pattern?
```{r}
result2 <- aggregate(ActivityData$steps, by = list(ActivityData$interval), FUN=mean)
colnames(result2)<- c("interval","steps")
plot(result2$interval, result2$steps, xlab="time interval", ylab="average steps", type='l', main="Average daily activity pattern")
```



## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
summary(rawData)

cat("\n as described above, total NA's in raw data is:", nrow(rawData[!complete.cases(rawData),]), "\n")

```

2. Devise a strategy for filling in all of the missing values in the dataset. Here we used the most simple strategy by filling the mean.
```{r}
rawData$steps.filled <- ifelse(is.na(rawData$steps),mean(rawData$steps, na.rm = TRUE),rawData$steps)
summary(rawData)
```

 3. a Histogram to show the result after imputing missing value.
```{r}
result.imputed <- aggregate(rawData$steps.filled, by = list(rawData$date), FUN=sum)
colnames(result.imputed)<- c("date","steps")
barplot(result.imputed$steps, xlab="date", ylab="total steps", space=0, names.arg = result.imputed$date, main = "Total number of steps taken each day (imputed)")
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
#chaning the time display else it would be Chinese charactor on my end.
Sys.setlocale(category = "LC_ALL", locale = "english")
rawData$day <- weekdays(as.Date(rawData$date), abbreviate = TRUE)
#rawData$week <- lubridate::week(as.Date(rawData$date))
rawData$isWeekday <- ifelse(rawData$day == "Sat" | rawData$day == "Sun", "weekday", "weekend")
result3 <- aggregate(rawData$steps, by = list(rawData$interval,rawData$isWeekday), FUN=mean, na.rm=TRUE)
colnames(result3)<- c("interval","isWeekday","steps.filled")

ggplot(data=result3, aes(x=interval, y=steps.filled, group=isWeekday, colour=isWeekday))+
       ggtitle("Ativity pattern (average step per 5 interval) color by weekdays / weekends") +
    geom_line() +
    geom_point()

```
